---
title: "PSY - Explorative Datenanalyse - v3"
author: "IMS-FHS"
date: "16 April 2019"
output: 
  html_document:
    toc: true
    toc_float: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, fig.width = 12, message = FALSE, cache = FALSE)
library(tidyverse)
library(ggplot2)
library(MASS)
library(haven)
library(GPArotation)
library(cluster)
library(klaR)
library(polycor)
library(mvtnorm)
library(car)
source("r/00-psy-import.R")

```


```{r echo=FALSE, include=TRUE, cache=TRUE}
# data preparation
# subset due to memory shortage
sample_size <- 3000
# sdf <- df[1:sample_size, ]
sdf <- df[sample(1:nrow(df), sample_size, replace = F), ]
# make factor variables out of booleans
sdf$iss_na <- NULL
sdf$ser_na <- NULL
col_ids <- grep("^iss_.+", colnames(sdf))
# only keep iss_* columns in sdf_num having at least one non-zero value
sdf_num <- sdf %>% dplyr::select(col_ids)
sdf_num <- sdf_num %>% mutate_each(as.numeric, 1:ncol(sdf_num))
# sdf_num <- sdf_num[, colSums(sdf_num != 0) > 0]

sdf_fact <- sdf_num %>% mutate_each(list(factor), 1:ncol(sdf_num))

# a) dissimilarity matrix by daisy, metric multidimensional scaling
# plot by dissimilarity matrix
diss_matrix <- daisy(sdf_fact)
coords <- cmdscale(diss_matrix)
```


# Übersicht

Grundlage sind Daten 2008 - 2018

## Infos
```{r}
cat("Anzahl Records:", nrow(sdf))
cat("Years: ", sort(unique(sdf$created_year)))
```

## Fälle nach Geschlecht

```{r}
barplot(table(sdf$gender)/nrow(sdf),  ylab = "%", main="Geschlecht")
```

```{r}
plot.new()
plot(coords, xlab = "Distanz x", ylab = "Distanz y")
legend("topright", 
  legend = c("female", "male"), 
  col = c("red", "blue"), 
  pch = c(19,19), 
  bty = "o", 
  pt.cex = 2, 
  cex = 1.2,  
  text.col = "black", 
  horiz = F , 
  inset = c(0.1, 0.1))
points(coords[which(sdf$gender == "F"),], col = "red")
points(coords[which(sdf$gender == "M"),], col = "blue")
```

## Fälle nach Sprache

```{r}
barplot(table(sdf$language_code)/nrow(sdf), ylab = "%", main="Sprache")
```


```{r}
plot.new()
plot(coords[which(sdf$language_code == "FR"),], col = "red", 
     xlab = "Distanz x", ylab = "Distanz y")
legend("topright", 
  legend = c("de", "it", "en", "fr"), 
  col = c("black", "blue", "green", "red"), 
  pch = rep(19,4), 
  bty = "o", 
  pt.cex = 2, 
  cex = 1.2,  
  text.col = "black", 
  horiz = F , 
  inset = c(0.1, 0.1))
points(coords[which(sdf$language_code == "EN"),], col = "green")
points(coords[which(sdf$language_code == "IT"),], col = "blue")
points(coords[which(sdf$language_code == "DE"),], col = "black")
```

## Fälle nach Schweregrad
```{r}
barplot(table(sdf$impact_on_work)/nrow(sdf), ylab = "%", main="Impact on Work")
```


```{r }
plot.new()
plot(coords[which(sdf$impact_on_work == "Normal"),], col = "black", 
     xlab = "Distanz x", ylab = "Distanz y")
legend("topright", 
  legend = c("Normal", "Satisfactory", "Impaired", "Severely Impaired", 
             "On Sick Leave/Absent Satisfactory"), 
  col = c("black", "green", "blue", "orange", "red"), 
  pch = rep(19,5), 
  bty = "o", 
  pt.cex = 2, 
  cex = 1.2,  
  text.col = "black", 
  horiz = F , 
  inset = c(0.1, 0.1))
points(coords[which(sdf$impact_on_work == "Satisfactory"),], col = "green")
points(coords[which(sdf$impact_on_work == "Impaired"),], col = "blue")
points(coords[which(sdf$impact_on_work == "Severely Impaired"),], col = "orange")
points(coords[which(sdf$impact_on_work == "On Sick Leave/Absent Satisfactory"),], col = "red")
```

## Fälle nach Clusterzugehörigkeit
```{r}
clust_num <- 5
plot.new()
cluster.results <- klaR::kmodes(sdf_fact, clust_num, iter.max = 10, weighted = FALSE )
summary(cluster.results)
plot(coords, col = cluster.results$cluster, 
     xlab = "Distanz x", ylab = "Distanz y")
legend("topright", 
  legend = 1:clust_num, 
  col = 1:clust_num, 
  pch = rep(19, clust_num), 
  bty = "o", 
  pt.cex = 2, 
  cex = 1.2,  
  text.col = "black", 
  horiz = F , 
  inset = c(0.1, 0.1))
# points(coords[which(sdf$impact_on_work == "Satisfactory"),], col = "green")
# points(coords[which(sdf$impact_on_work == "Impaired"),], col = "blue")


```



## Exploratory Factor Analysis 
Quelle Siehe: 
- https://www.statmethods.net/advstats/factor.html
- http://www.psychologie.uzh.ch/dam/jcr:ffffffff-852f-247f-ffff-ffff99c06567/explorative_faktorenanalyse_mit_r_cswerner.pdf

### Scree Plot

Scree plot
A scree plot displays the proportion of the total variation in a dataset that is explained by each of the components in a principle component analysis. It helps you to identify how many of the components are needed to summarise the data.


```{r}

screeplot(princomp(sdf_num))

# fit <- factanal(sdf_num, 3, rotation="varimax")
# print(fit, digits=2, cutoff=.3, sort=TRUE)
# # plot factor 1 by factor 2 
# load <- fit$loadings[,1:2] 
# plot(load,type="n") # set up plot 
# text(load,labels=names(mydata),cex=.7) # add variable names


```

Basierend auf diesem Scree-Plot kann davon ausgegangen werden, dass zur Erklärung des grössten Teils der Varianz durch den Einbezug von 4 Hauptkomponenten erklärt werden kann.


```{r}
num_comp_to_extract <- 4
# berechnung korrelationsmatrix
# cor_mat <- cor(sdf_num)
# diag(cor_mat) <- psych::smc(sdf_num)
# cor_mat <- polycor::hetcor(sdf_fact, ML = TRUE)
# plot(eigen(cor_mat)$values, ylab = "Eigenvalue", xlab = "Factor", 
#      type = "b", ylim = c(0,1.5))

# fit_fa <- psych::fa(r = cor_mat, nfactors = num_comp_to_extract, rotate = "varimax")
# summary(fit_fa)

# fit <- psych::principal(r = cor_mat, nfactors = num_comp_to_extract, rotate = "promax")

```


